---
title: "market"
author: "Crystal Tse"
date: "4/20/2020"
output: md_document
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, cache = TRUE,autodep = TRUE ,cache.comments = FALSE, message = FALSE, warning = FALSE)

# reordered with the PCA in front
library(mosaic)
library(tidyverse)
library(ggplot2)
library(LICORS)  # for kmeans++
library(foreach)
library(reshape2)
library(knitr)
library(kableExtra)
library(cluster)
library(factoextra)
library(lattice)
library(gridExtra)
```

# Market Segmentation

To analyze the potential market for NutrientH2O, a sample of followers' tweets over a seven-day period was analyzed and placed into 36 broad categories for subject matter. This service was performed by contracted workers of Amazon's Mechanical Turk service: human annotators that are prone to error and subjectivity. Some categories included were politics, food, and school. However, there were also categories such as spam and adult, whose observations were cut out of the data to focus on the most pertinent individuals. But certain ambiguities still remain, such as miscellaneous categories such as "chatter" and "uncategorized," which leads to even more noise in the data.

## PCA Attempt
Principal component analysis (PCA) seems like it would be a good method to analyze potential markets, since it is adept at handling noisy data and can accommodate ambiguity, like tweets that lie in multiple categories, unlike clusters, which strictly group individual points into sections.

```{r echo=FALSE}
mkt = read.csv("social_marketing.csv")
mkt <- subset(mkt, !(spam > 0 | adult > 0))

mkt = subset(mkt, select = -c(spam, adult) )
# Center and scale the data
mkt = scale(mkt[-1], center=TRUE, scale=TRUE)

#remove the categorical variables and scale
# mkt <- as.data.frame(scale(mkt[-1]))

# Extract the centers and scales from the rescaled data (which are named attributes)
mu = attr(mkt,"scaled:center")
sigma = attr(mkt,"scaled:scale")

mkt_long <- reshape2::melt(mkt)  # convert matrix to long dataframe
mkt <- spread(mkt_long, Var2, value)# convert long dataframe to wide
```

From the below summary, it appears as though 15 principal components are necessary to explain about 75% of the variation in the data. This indicates that the data is not easily summarized; the subject matter variables are not differentiated enough to generate principal components that explain a large percentage of the variation in the data. For example, food, health_nutrition, and cooking could all be overlapping but lead to small fluctuations that lead to overfitting and more principal components than necessary.

```{r echo = FALSE}
# Now look at PCA of the (average) survey responses.  
# This is a common way to treat survey data
tweets <- mkt
PCAtweet = prcomp(tweets, scale=TRUE)
summary(PCAtweet)

```

The scree plot below confirms that PCA is not a good method for summarizing the data. The "elbow" in this plot that shows the dropping-off point, or diminishing marginal utility, of adding more principal components seems to be after only one principal component. In reality, this is not reasonable since this component only explains 13% of the variation in the data. Fifteen principal components are needed to explain 75% of the data, but this makes market segmentation more confusing and convoluted.

There are too many principal components needed to explain the variation in the data. Each principal component pulls apart a market segment, and in doing so, takes us away from the big picture audience and plunges us too much in the messy details.

```{r echo = FALSE}
fviz_screeplot(PCAtweet, addlabels=TRUE)

```

## K-Means Clustering Attempt
Instead, K-means clustering might generate more interpretable and concise results.

```{r echo = FALSE}
# try K-means++ clustering

# choose the optimal K
fviz_nbclust(mkt, kmeans, method="wss")
gap_stat <- clusGap(mkt, FUN = kmeans, nstart = 25,
                    K.max = 10, B = 50)
fviz_gap_stat(gap_stat)
```

Although the Gap statistic indicates that the best number of clusters is supposedly 1, the value for k after which there is a dip in the calculated Gap statistic, this is not a clear solution. The Gap statistic values are all close to each other. In this case, the most intuitive number of clusters is 3 clusters.


```{r echo = FALSE}

# Run k-means plus plus.
clust2 = kmeanspp(tweets[-1], k=3, nstart=25)  # edit with optimal_k

c1 = clust2$center[1,]*sigma + mu
c2 = clust2$center[2,]*sigma + mu
c3 = clust2$center[3,]*sigma + mu
```

Below are the top 5 categories for each cluster, or group of individuals that are closest together in their tweets across all the 37 different topics that were measured. Although "chatter" appears for all 3 clusters, it is not a very significant measure that can be targeted since it is somewhat of a "miscellaneous" category and is fraught with noise. However, it is a gauge of how relatively active each cluster is, and was therefore kept in these charts.

The value "X" shown is the number of standard deviations above the average each cluster tweets about a certain topic.

```{r echo = FALSE}
# show the variables that are above 4 or the top 5 variables that describe each cluster
(c1[order(c1, decreasing = TRUE)][1:5])%>% kable(caption = "Cluster 1") %>%  kable_styling(bootstrap_options = "striped", full_width = F)
```

```{r echo = FALSE}
(c2[order(c2, decreasing = TRUE)][1:5])%>% kable(caption = "Cluster 2") %>% kable_styling(bootstrap_options = "striped", full_width = F)
```

```{r echo = FALSE}
(c3[order(c3, decreasing = TRUE)][1:5]) %>% kable(caption = "Cluster 3") %>% kable_styling( bootstrap_options = "striped", full_width = F)
```


It seems as though one cluster has relatively inactive members that do not skew strongly towards certain topics in their tweets; these do not present a strong demographic to market towards.

However, the other two clusters present clearer trends. One of these clusters represents a group of followers that is very interested in health and nutrition, sharing photos, cooking, and politics. This suggests that expanding to a platform that showcases more photos (such as Instagram) and emphasizing the palatability and picture-worthy aspects of NutrientH2O products would be highly effective.

Another cluster shows that many followers of Nutrient H2O are devoted to sports, religion, food, and parenting. This suggests that targeting parents during certain sports seasons may be effective, for example.

## A Graphical Analysis with Clusters

Below are some plots that demonstrate the different market segments in NutrientH2O's followers, focusing on variables that had a strong showing in the 3 clusters.

The plot below shows that photo sharing and sports fanaticism are not strongly correlated, so targeting the intersection of those two groups is not recommended.

```{r echo = FALSE}
# A few plots with cluster membership shown
# sports_photos = 
  ggplot(data = mkt,
       aes(x = sports_fandom, y = photo_sharing, color = factor(clust2$cluster))) +
  geom_point(position = "jitter")+
  ggtitle("Sports and Photo-sharing Are Not Correlated")+
  labs(color = "Cluster")
```

However, sports fanaticism and parenting seem to be highly correlated. It is worth investigating parents that are into sports as a potential market segment.

```{r echo=FALSE}
# sports_parents =
  ggplot(data = mkt, 
       aes(x = sports_fandom, y = parenting, color = factor(clust2$cluster))) +
  geom_point(position = "jitter")+
  ggtitle("Sports and Parenting are Highly Correlated")+
  labs(color = "Cluster")
```

Although these clusters are not particularly distinct in the scatterplot below, it shows that health_nutrition and personal_fitness are highly correlated. This is an example of how some of the variables tracking subject matter measure very similar things.

```{r echo = FALSE}
# large portion of their market is focused on health/nutrition and personal fitness
# health_fitness =
  ggplot(data = mkt, 
       aes(x = health_nutrition, y = personal_fitness, color = factor(clust2$cluster))) +
  geom_point(position = "jitter")+
  ggtitle("Health/Nutrition and Personal Fitness are Highly Correlated")+
  labs(color = "Cluster")
```

Sports fanaticism and religion are also highly correlated. However, in promotions, it might be wise to target sports more than religion, as it seems there is a greater concentration along the sports axis than the religion axis for individuals that are not strongly into both topics.

```{r echo=FALSE}
# sports fanaticism and religion correlated
# sports_religion = 
  ggplot(data = mkt,
       aes(x = sports_fandom, y = religion, color = factor(clust2$cluster))) +
  geom_point(position = "jitter")+
  ggtitle("Sports and Religion are Highly Correlated")+
  labs(color = "Cluster")
```

Religion and food are also highly correlated. There appears to be a main cluster of people that are both religious and into food; however, in promotions, it would be wise to err on the side of featuring food because there seems to be a greater concentration along the food axis in a secondary cluster.

```{r echo = FALSE}
# religion and food seem highly correlated, can have holiday specials or something
# religion_food = 
  ggplot(data = mkt, 
       aes(x =religion, y = food, color = factor(clust2$cluster))) +
  geom_point(position = "jitter")+
  ggtitle("Religion and Food are Highly Correlated")+
  labs(color = "Cluster")
```

```{r echo = FALSE}
# food and politics are not correlated, don't try to tackle both
# food_polit = 
  ggplot(data = mkt,
       aes(x = politics, y = food, color = factor(clust2$cluster))) +
  geom_point(position = "jitter")+
  ggtitle("Politics and Food are Not Correlated")+
  labs(color = "Cluster")
```

Here, food and parenting also seem to be highly correlated.

```{r echo=FALSE}
# food and parenting seem to be pretty correlated, they cluster together
# can target with ads
# food_parent = 
  ggplot(data = mkt,
       aes(x = food, y = parenting, color = factor(clust2$cluster))) +
  geom_point(position = "jitter") +
  ggtitle("Food and Parenting are Highly Correlated")+
  labs(color = "Cluster")
```

However, targeting these parents with pictures of food or visual ads in general would not necessarily be an effective strategy.

```{r echo = FALSE}
# but parents aren't necessarily attracted to photos, don't need to target through visual ads for example
# parent_photo = 
  ggplot(data = mkt, 
       aes(x =parenting, y = photo_sharing, color = factor(clust2$cluster))) +
  geom_point(position = "jitter")+
  ggtitle("Parenting and Photo-Sharing are Not Correlated")+
  labs(color = "Cluster")
```


## Conclusion

Overall, clustering was determined to be a more intuitive method than PCA for determining market segments. The different groups were easier to visualize, versus PCA which condensed the data into 15 different principal components that separated out the overall audience into too many subcategories.

NutrientH2O's Twitter followers can be broken into 3 clusters. One cluster seems to be relatively inactive, suggesting disengagement and an audience that is not easily targeted because of a lack of engagement and demonstrated preferences. The other two clusters show more apparent trends. 

Food, parenting, religion, and sports are all shown to be key interests of NutrientH2O's active follower base. However, certain strategies may not be as effective in targeting these audiences. For instance, parents are not keen on photo-sharing and may not be as receptive to solely visual advertisements.

